---
title: "Project 2 Description"
author: "Team 7"
date: "`r Sys.Date()`"
output: html_document
---

***


## Abstract

Summarize analysis - to do after analysis


## Introduction

The dataset from the World Health Organization (WHO) consists of daily reported counts of cases and deaths due to COVID-19. We also referenced the Trips by Distance dataset from the Bureau of Transportation Statistics for information on the number of trips taken by distance of the nation during the pandemic. In this data analysis, we are primarily interested in the following questions: if there are variations between different distances for trips taken outside of the home and what is the relationship between the different distances of trips and COVID-19 cases and deaths. The data analysis will distinguish if trip distances affect COVID-19 cases. This will indicate if increased travel contributes to a rise of cases and deaths, primarily due to community transmission.

# Background

This data analysis utilizes the two following datasets: the WHO COVID-19 Dashboard and the Monthly Transportation Statistics.

The WHO has consistently collected data from around the world during the COVID-19 pandemic. The WHO dataset reports daily confirmed counts by countries, territories, and areas. Counts starting from December 31, 2019 to March 21, 2020 were taken from International Health Regulations (IHR) and thereafter, the confirmed counts are continuously compiled from global data on WHO dashboards of the headquarters and different regions. Daily counts may vary between countries, territories, and areas due to different sources, cut-off times, case detections, etc.

The WHO COVID-19 dataset used in the analysis contains daily updated counts of cases and deaths with the following variables: the reporting date, country code, country, WHO region, new cases count, cumulative cases to date, new deaths count, and cumulative deaths to date.

The Trips by Distance dataset compiles statistics of transportation, including the distance people travel when they are not quarantining during the COVID-19 pandemic and stay-at-home order. The Bureau of Transportation Statistics has reported data to date of mobility statistics. The data is gathered and merged from a mobile device data panel that takes distances traveled by people. The Bureau considers temporal frequency and spatial accuracy for locations as well as temporal coverage and representativeness for devices. All personal mobile data and locations have been excluded from sources for confidentiality. These trips include all types of transportation, including but not limited to airlines, transit, trains, and cars.

The dataset contains daily updated statistics with variables, such as the national count of the population staying at home, the population not staying at home, and the count of trips of various distances from less than one mile to over 500 miles, which are the primary variables of this data analysis.


```{r include = F, echo = FALSE}
library(dplyr)
library(tibble)
library(tidyverse)
library(ggplot2)
library(car)
library(stringr)
library(readr)
knitr::opts_chunk$set(fig.pos = 'H')
```

```{r, message = FALSE, results = FALSE, warning = FALSE}
tripsbydistance <- read_csv("Trips_by_Distance.csv") 
names(tripsbydistance)<-str_replace_all(names(tripsbydistance), c(" " = "_"))

WHO_COVID_19_global_data <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
us_covid_data <- filter(WHO_COVID_19_global_data, Country %in% "United States of America")

mergedata = merge(tripsbydistance ,us_covid_data, by.x = "Date", by.y = "Date_reported")
newdata <- mergedata[-c(2:9, 20:21)]
head(newdata)

casemortalityrate <- newdata$New_deaths / newdata$Cumulative_cases
casemortalityrate
```

Below is an interactive plot to show the changes in new cases and deaths across time, starting from when cases and deaths started to be reported. The delay of confirmed cases is attributed to the delay of action and discovery of the coronavirus. 

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(plotly)
newdata %>% 
  filter(Date>= "2020-03-01", Date<= "2020-11-30") %>% 
  group_by(Date,WHO_region) %>%   summarize(deaths = sum(New_deaths),
            cases = sum(New_cases)) %>% 
  mutate(Date = Date- as.Date("2020-03-01")) %>%
  plot_ly(
    x= ~cases,
    y= ~deaths,
    frame = ~Date,
    type = 'scatter',
    mode = 'markers',
    showlegend = T
  )
```

Another plot is included below to show the changes in new cases over time. This can reflect trends of society with seasons as well as policy changes. It is observed that there was a substantial increase during the summer months. This shows there was an increase community transmission due to society leaving their homes more often as well as the changes in policies to improve the economic environment with relaxing certain restrictions for businesses of the stay-at-home order. It is also observed that there was a very large increase during the holiday season, ranging from November to January. Families are more likely to have larger group sizes and more likely to travel to visit family and friends.

```{r, echo = FALSE}
fig2 <- plot_ly(newdata, x = ~Date, y = ~New_cases, name = 'trace 0', type = 'scatter', mode = 'lines') %>%   layout(title = "New Cases vs Date",
         xaxis = list(title = "Date"),
         yaxis = list (title = "New Cases"))
fig2
```

The plot below is of new deaths versus new cases. Here, we can observe that although when new cases start to rise, the number of new deaths also starts to rise. The points seem to follow a linear, positive relationship.

```{r, echo = FALSE}
fig3 <- newdata %>% 
  filter(Date>= "2020-03-01", Date<= "2020-11-30") %>% 
  ggplot(aes(x=New_cases,y=New_deaths)) +
  geom_point()+
  geom_text(aes(label = " "), hjust=0, vjust=0) +
  ggtitle("New Deaths vs New Cases") + 
  geom_smooth(method='lm', formula= y~x) + 
  xlab("New Cases") + ylab("New Deaths")
fig3
```

```{r, echo = FALSE}
fig3 <- newdata %>% 
  filter(Date>= "2020-03-01", Date<= "2020-11-30") %>% 
  ggplot(aes(x=Date,y=casemortalityrate)) +
  geom_point()+
  geom_text(aes(label = " "), hjust=0, vjust=0) +
  ggtitle("New Deaths vs New Cases") + 
  geom_smooth(method='lm', formula= y~x) + 
  xlab("New Cases") + ylab("New Deaths")
fig3
```


## Descriptive Analysis

```{r}
linear.model1 = lm(New_cases~
`Number_of_Trips_<1`+ `Number_of_Trips_1-3` + `Number_of_Trips_3-5` + `Number_of_Trips_5-10` + `Number_of_Trips_10-25` + `Number_of_Trips_25-50`+ `Number_of_Trips_50-100` + `Number_of_Trips_100-250` + `Number_of_Trips_250-500` + `Number_of_Trips_>=500`, data=newdata)
summary(linear.model1)
```


## Inferential Analysis 

Propose an appropriate model to answer the questions of interest. 
a. Explain your notations. 
b. State assumptions for your model. 
c. Explain why your model is appropriate for this task. 

Fit the proposed model in (4) and explain your results.

## Sensitivity Analysis

Conduct model diagnostics

## Discussion

Conclude your analysis with a discussion of your findings and caveats of your approach.

## References

WHO coronavirus disease (COVID-19) dashboard. Geneva: World Health Organization, 2020. Available online: https://covid19.who.int/ 




