---
title: "Project 2 Description"
author: "Team 7"
date: "`r Sys.Date()`"
output: html_document
---

***


## Abstract

Summarize analysis - to do after analysis


## Introduction

Introduce the dataset and the questions of interest.

# Background

Review the background of the coronavirus.

```{r include=F}
library(dplyr)
library(tibble)
library(tidyverse)
library(ggplot2)
library(car)
library(stringr)
library(readr)
knitr::opts_chunk$set(fig.pos = 'H')

```

```{r  echo=F}
monthly_flights = read_csv("Monthly_Transportation_Statistics.csv") 
new_data <- tidyr::separate(monthly_flights, Date, c("date", "time"), sep = " ")
new_data

new_flight_data <- new_data[,c(1:3,7,9,45:47,92,98,100,129,133)]
new_flight_data


WHO_COVID_19_global_data <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
us_covid_data <- filter(WHO_COVID_19_global_data, Country %in% "United States of America") 
us_covid_data
```




## Descriptive Analysis

Explore this dataset and generate summary statistics that you find informative, and explain your findings. <span style="color:red"> The summary statistics should include at least time, number of cases, number of death, case-mortality rate.</span>

```{r}
library(plotly)
us_covid_data %>% 
  filter(Date_reported>= "2019-11-01", Date_reported<= "2020-11-30") %>% 
  group_by(Date_reported,WHO_region) %>%   summarize(deaths = sum(New_deaths),
            cases = sum(New_cases)) %>% 
  mutate(Days_11.19_11.20 = Date_reported- as.Date("2019-11-01")) %>%
  plot_ly(
    x= ~cases,
    y= ~deaths,
    frame = ~Days_11.19_11.20,
    text=~WHO_region,
    hoverinfo="WHO_region",
    color=~WHO_region,
    type = 'scatter',
    mode = 'markers',
    showlegend = T
  )
```

```{r}
fig.spaghetti.1 <- us_covid_data %>% 
  filter(Date_reported>= "2019-11-01", Date_reported<= "2020-11-30") %>% 
  mutate(Date=as.Date(Date_reported)) %>%
  ggplot(aes(x=Date,y=New_cases,by=Country)) +
  geom_line(aes(color=Country)) +
  theme(legend.position ='none')
fig.spaghetti.1   
```




## Inferential Analysis 

Propose an appropriate model to answer the questions of interest. 
a. Explain your notations. 
b. State assumptions for your model. 
c. Explain why your model is appropriate for this task. 

Fit the proposed model in (4) and explain your results.

## Sensitivity Analysis

Conduct model diagnostics

## Discussion

Conclude your analysis with a discussion of your findings and caveats of your approach.




